Dynamic Stochastic Approximation Monte Carlo \jpsays{DSA for stochastic optimization (not Monte Carlo) is actually a thing (1965, 2017) although unrealated to what we are doing}

### 2018-07-27 papers/histogram/sad.tex
added section on binning and made a plotting fix

\begin{document}
\title{Dynamic Stochastic Approximation Monte Carlo}

@@

\davidsays{We should discuss continuous energies and binning somewhere
  here...}
\jpsays{I put some text about this but it was under a sad subsection also
may need reworked and/or edited.}

@@

Say something about detailed balance here?...

@@

While the current algorithm is highly efficient, a weighting function $\eta_t$
is introduced to ensure that the system samples all states
evenly~\cite{mahynski2017multivariable, blanco2016effect}.
As for the Flat-histogram approach, the weighting function is
set inversely proportional to the current macrostate probability.
\begin{align}
  \eta(I) = - \ln(D(I))
\end{align}
The probability is defined in terms of this biasing function:
\begin{align}
  p(I\rightarrow J) = \min\bigg[1,\frac{\eta(J)}
  {\eta(I)} e^{\beta (E(I)-E(J))}\bigg]
\end{align}
The TMMC algorithm consists of recording the data in a collection matrix
(the unormalized transition histogram matrix).
\begin{align}
  C(I \rightarrow J) = C(I \rightarrow J) + p(I \rightarrow J)\\
  C(I \rightarrow I) = C(I \rightarrow I) +(1 - p(I \rightarrow J))
\end{align}
If the energy state has not been changed, then $C$ is still incremented by
unity. Fitzgerald et al.~\cite{fitzgerald2000monte} showed that defining
the third step this way results in a uniform improvement over using
histogram estimators to determine the particle number probability
distribution (PNPD). The algorithm is highly efficient due to the ability
to bias the simulation while keeping the criteria used to update the
collection matrix $C(I\rightarrow J)$.
}

>>>>

\begin{document}
\title{Dynamic Stochastic Approximation Monte Carlo
\jpsays{DSA for stochastic optimization (not Monte Carlo) is actually a 
thing (1965, 2017) although unrealated to what we are doing}
}

@@

{\color{red} Both weight-based and transition matrix methods have
calculate the density of states $D(E)$ for discrete energy
levels~\cite{haber2018performance}. Energy binning becomes an important
consideration for complex continous systems for this reason. Energy
bins are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}. Coarse binning yields fewer
energy states to explore but more time spent in each individual bin.
Fine binning yields more energy states to explore with less time spent
in each bin. Some methods such as AdaWL~\cite{koh2013dynamically}
employ a tunable mechanism for controlling the binning for low entropic
states in order to ensure the exploration of all energies.
}

@@

Say something about collecting statistics vs WL?
Say something about detailed balance here?...

@@

While the current algorithm is highly efficient, a weighting function $\eta(I)$
is introduced to ensure that the system samples all states
evenly~\cite{mahynski2017multivariable, blanco2016effect}.
As for the Flat-histogram approach, the weighting function is
set inversely proportional to the current macrostate probability.
\begin{align}
  \eta(I) = - \ln(D(I))
\end{align}
The probability is defined in terms of this biasing function:
\begin{align}
  p(I\rightarrow J) = \min\bigg[1,\frac{\eta(J)}
  {\eta(I)} e^{\beta (E(I)-E(J))}\bigg]
\end{align}
The TMMC algorithm consists of recording the data in a collection matrix
(the unormalized transition histogram matrix).
\begin{align}
  C(I \rightarrow J) = C(I \rightarrow J) + p(I \rightarrow J)\\
  C(I \rightarrow I) = C(I \rightarrow I) +(1 - p(I \rightarrow J))
\end{align}
If the energy state has not been changed, then $C$ is still incremented by
unity. Fitzgerald et al.~\cite{fitzgerald2000monte} showed that defining
the third step this way results in a uniform improvement over using
histogram estimators to determine the particle number probability
distribution (PNPD). The algorithm is highly efficient due to the ability
to bias the simulation while keeping the criteria used to update the
collection matrix $C(I\rightarrow J)$.
}
